---
title: "Data Wizards (Group 4) Project 2"
author:
- Di Chen
- Mai Castellano
- Tyler Kussee
- Spencer (Hutchison) Yang
output: pdf_document
geometry: margin=0.7in
---
\vspace{-5truemm}

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
##An if statement for checking if a package is installed 

if (!require(tidycensus)) {
  install.packages("tidycensus")
}
if (!require(tidyverse)) {
  install.packages("tidyverse")
}
if (!require(dplyr)) {
  install.packages("dplyr")
}
if (!require(ggplot2)) {
  install.packages("ggplot2")
}

# Reading in the csv
train <- read.csv('train.csv')
train <- train %>% select(-"DISBURSAL_DATE")
train$DATE_OF_BIRTH <- as.Date(train$DATE_OF_BIRTH, format = "%d-%m-%Y")
train$AGE <- as.Date('01-01-2019', format = "%d-%m-%Y") - train$DATE_OF_BIRTH
train$AGE <- as.integer(floor(train$AGE / 365.25))
train <- train %>% select(-"DATE_OF_BIRTH")
# Calculate strange data fields
train$acctyr <- as.numeric(gsub("yrs.*", "", train$AVERAGE_ACCT_AGE))
train$acctmo <- as.numeric(gsub(".*yrs|mon", "", train$AVERAGE_ACCT_AGE))
train$crdtyr <- as.numeric(gsub("yrs.*", "", train$CREDIT_HISTORY_LENGTH))
train$crdtmo <- as.numeric(gsub(".*yrs|mon", "", train$CREDIT_HISTORY_LENGTH))
# Replace strange data fields
train$AVERAGE_ACCT_AGE <- round(train$acctyr + train$acctmo / 12, 2)
train$CREDIT_HISTORY_LENGTH <- round(train$crdtyr + train$crdtmo / 12, 2)
# Remove calc fields
train <- train %>% select(-acctyr, -acctmo, -crdtyr, -crdtmo)
# Create emloyment dummy fields
train$SELF_EMPLOYED <- ifelse(train$EMPLOYMENT_TYPE == "Self employed", 1, 0)
train$SALARIED <- ifelse(train$EMPLOYMENT_TYPE == "Salaried", 1, 0)
train$NULL_EMPLOYMENT <- ifelse(is.na(train$EMPLOYMENT_TYPE), 1, 0)
# Remove employment_type
train <- train %>% select(-EMPLOYMENT_TYPE)
# Pull CNS score letter grade
train$PERFORM_CNS_SCORE_DESCRIPTION <- as.factor(substr(train$PERFORM_CNS_SCORE_DESCRIPTION,
                                              1, 1))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#Load libraries
library(tidycensus)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Introduction

In our pursuit of statistical inquiry, we have chosen to explore the Vehicle Loan Default dataset, comprising approximately 41 columns, with one designated as the response variable. Encompassing diverse information, the dataset delves into loanee details, including date of birth, employment type, and credit score, alongside loan-related specifics such as disbursal details and loan-to-value ratios. The dataset presents challenges, notably in the form of odd date and time length columns, requiring standardization and transformation into comprehensible formats conducive to model development.

We want to discover the most influential explanatory variables driving loan default, and their impact within the dataset. We also want to find the optimal modeling approach for harnessing the training data, evaluating various methodologies to identify the most effective. Ultimately, our investigation extends to which among them best identifies the underlying dynamics of vehicle loan default prediction.


## Obtain/Scrub the data

## Explore the data

## Model the data

## Interpret the data

## Obstacles

## Conclusion

## Appendix
